---
title: "Exercise 2.18"
author: "Student Name"
date: "2023-02-08"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


```{r shoppers_rating}
library(ggplot2)

shoppers_rating = c(26, 33, 33, 29, 34, 32, 22, 25, 27, 33, 34, 31, 20, 24, 30, 31, 28, 30, 31, 32, 32, 31, 28,30, 31,33, 32, 32, 34, 30, 33, 32, 25, 33, 28, 29, 27, 32, 32, 33, 31, 32, 30, 33, 35, 31, 33, 29, 
34,27, 34, 32, 28, 26, 29, 31, 29, 33, 30,32)
```
# a Find the number of classes that should be used to construct a frequency distribution and histogram for the bottle design ratings.

```{r}
length(unique(shoppers_rating))
```
# b If we round up to the nearest whole rating point, show that we should employ a class length equal to 3.

```{r}
data_rounded <- round(shoppers_rating)
data_range <- range(data_rounded)
class_length <- 3
number_of_classes <- ceiling((data_range[2] - data_range[1]) / class_length)
hist(data_rounded, breaks = seq(data_range[1], data_range[2], length.out = number_of_classes + 1), col = "lightblue")
```
# c Define the nonoverlapping classes for a frequency distribution.

Nonoverlapping classes for a frequency distribution refer to the ranges or intervals that are created to categorize the data. The classes are defined in such a way that they do not overlap or intersect each other, and each data point belongs to only one class.

# d Tally the number of ratings in each class and develop a frequency distribution.

```{r}
classes <- cut(shoppers_rating, breaks=seq(20, 35, by=3), include.lowest=TRUE)
levels(classes) <- c("20-23", "23-26", "26-29", "29-32", "32-35")


# Tally the number of ratings in each class


table(classes)
```
```{r}
# frequency distribution
table <- as.data.frame(table(classes))

ggplot(data = table, aes(x = classes, y = Freq, fill=classes)) + 
  geom_bar(stat = "identity") +
  labs(x = "classes",
       y = "Average Rating") +
  scale_fill_hue (c = 50) + theme (legend.position = "none") +
  ggtitle("Histogram of Each Class")

```

# e Draw the frequency histogram for the ratings data, and describe the distribution shape.

```{r}
# Create a histogram
hist <- ggplot(data=data.frame(shoppers_rating), aes(x=shoppers_rating)) +
  geom_histogram(binwidth=0.5, fill="lightblue", color="black")

# Add a frequency polygon to the histogram
hist + geom_density(fill="red", alpha=0.5)
```
The distribution shape frequency histogram for ratings data that has negative skewness and is asymmetrical indicates that the majority of the ratings data is concentrated on the right side of the histogram.
