---
title: "Lab 7"
author: "Neel Patel"
date: "2022-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#On your own

```{r names}
download.file("http://www.openintro.org/stat/data/nc.RData", destfile = "nc.RData")
load("nc.RData")
names(nc)
```

```{r}

# The data summary

summary(nc)
```



#1.Calculate a 95% confidence interval for the average length of pregnancies (weeks) and interpret it in context. Note that since you're doing inference on a single population parameter, there is no explanatory variable, so you can omit the x variable from the function.

```{r}
# The argument type = "ci" refers "Confidence interval"
inference(y = nc$weeks, est = "mean", 
          type = "ci", null = 0, alternative = "twosided",
          method = "theoretical")
```

There is 95 % Confidence interval that the average length of pregnancy in weeks is between 38.1528 ,and 38.5165.

#2.Calculate a new confidence interval for the same parameter at the 90% confidence level. You can change the confidence level by adding a new argument to the function: conflevel = 0.90.

```{r}

# ADD conflevel = 0.90

inference(y = nc$weeks, est = "mean", 
          type = "ci", null = 0, alternative = "twosided",
          method = "theoretical", conflevel = 0.90)
```


#3.Conduct a hypothesis test evaluating whether the average weight gained by younger mothers is different than the average weight gained by mature mothers.

```{r}
inference (nc$weight, nc$mature, 
           est = "mean", type = "ht", 
           null = 0, alternative = "twosided", method = "theoretical")

```


#4.Now, a non-inference task: Determine the age cutoff for younger and mature mothers. Use a method of your choice, and explain how your method works.

```{r}
#Determine age cutoff for younger mothers    
print(paste(min(subset(nc, 
                       mature == 'younger mom')$mage, 
                na.rm = TRUE),
            " is the minimum cut off age for younger mother"))


print(paste(max(subset(nc, 
                       mature == 'younger mom')$mage, 
                na.rm = TRUE),
            " is the maximum cut off age for younger mother"))
```



```{r}

#Determine age cutoff for mature mothers    

print(paste(min(subset(nc, 
                       mature == 'mature mom')$mage, 
                na.rm = TRUE),
            " is the minimum cut off age for mature mother"))


print(paste(max(subset(nc, 
                       mature == 'mature mom')$mage, 
                na.rm = TRUE),
            " is the maximum cut off age for mature mother"))
```
The Method works by sub-setting the dataset based on the maturity status of mother, i first subset a dataset for younger mother and select the mother the mothers age column. From this column i select both the minumum and the maximum values for both younger and mature mothers as their age cut offs, as shown above. Lastly i print the results.

#5.Pick a pair of numerical and categorical variables and come up with a research question evaluating the relationship between these variables. Formulate the question in a way that it can be answered using a hypothesis test and/or a confidence interval. Answer your question using the inference function, report the statistical results, and also provide an explanation in plain language.

Variables chosen are the length of pregnancy in weeks and whether mom is white or not white. which are weeks and white mom.

Question: Is there a difference in the length of pregnancy in weeks for mothers if she is is white or not white?

The formulated Null hypothesis states that there is a difference in the pregnancy period between a white and a non white mother.

```{r}

inference (y= nc$weeks, x= nc$whitemom, est = "mean",
           type = "ht", null = 0, 
           alternative = "twosided", method = "theoretical")
```
We observe from the statistical analysis that the p value is 0.0076 which is less than 0.05. This shows that the test is statistically significant and we therefore reject the Null hypothesis that there is a difference in the pregnancy period between a white and a non white mother.

