---
title: "Question 3"
author: "Neel Patel"
date: "2022-12-17"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##(1) ANOVA:

Capture an ANOVA model on the iris dataset. Set the dependent variable to 'Species'.
Capture a summary. 
Provide an interpretation of the results in your own words.  Support your response with results captured from running ANOVA.

```{r iris}
## Load the necessary libraries
library(ez)
library(MOTE)
library(ggplot2)
library(pwr)

##data
iris = read.csv("iris_exams.csv")


# Accuracy
str(iris)
unique(iris$Species)
iris$Species = as.factor(iris$Species)
str(iris$Species)
```

Outliers
```{r}
# Outliers

#  summary of your mahal scores.
mahal = mahalanobis(iris[ , 3:6], colMeans(iris[ , 3:6], na.rm = TRUE),
                    cov(iris[ , 3:6], use = "pairwise.complete.obs"))

summary(mahal)

# the df for your Mahalanobis cutoff?
cutoff = qchisq(1-.001, ncol(iris[, 3:6]))
cutoff

## df and cutoff
ncol(iris[,3:6])
#  outliers
summary(mahal < cutoff)

#  	Delete the outliers.
noout = subset(iris, mahal < cutoff)
str(noout)
```

```{r}
noout$partno = 1:nrow(noout)
options(scipen = 999)
noout$Species = as.numeric(noout$Species)
noout$partno = as.factor(noout$partno)
noout$Sepal.Length = as.numeric(noout$Sepal.Length)

ezANOVA(data = noout,
        dv = Species,
        between = Sepal.Length,
        wid = partno,
        type = 3, 
        detailed = T)

```
Levene’s test is significant, F(1,296)=123.9346, p=0.0000000000000000000000028040797422941528. The results suggest that there is no significant difference between the Species and Sepal Length in the iris dataset. We can also use the one-way ANOVA to test if there is a significant difference between Sepal Length and Species in iris dataset. The results show that there is no significant difference between the Species and Sepal Length in the iris dataset


