---
title: "Question 4"
author: "Neel Patel"
date: "2022-12-17"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Provide at least the following in the report for full credit: 

- Null Hypothesis H0 : Sepal.Length has no effect on Species (Setosa & Versicolor Only). That is to say that the difference between the observed Sepal.Length values for various Species are not statistically different

- Alternate Hypothese Ha : Sepal.Length has some effect on Species (Setosa & Versicolor Only). That is to say that the difference between the observed Sepal.Length values for various Species are in fact different from each other.

(1) The t-test :

Capture a t-test to compare the  Sepal.Length of the Species (Setosa & Versicolor Only) from the iris dataset. 
Capture a summary. 
Provide an interpretation of the results in your own words.  Support your response with results captured from performing a t-test.

```{r}
##data
iris = read.csv("iris_exams.csv")


# Accuracy
str(iris)
unique(iris$Species)


# Subset Species for Setosa & Versicolor Only

iris1 = iris[iris$Species %in% c('setosa','versicolor'),]
unique(iris1$Species)
```

```{r}
# Outliers

#  summary of your mahal scores.
mahal = mahalanobis(iris1[ , 3:6], colMeans(iris1[ , 3:6], na.rm = TRUE),
                    cov(iris1[ , 3:6], use = "pairwise.complete.obs"))

summary(mahal)

# the df for your Mahalanobis cutoff?
cutoff = qchisq(1-.001, ncol(iris1[, 3:6]))
cutoff

## df and cutoff
ncol(iris1[,3:6])
#  outliers
summary(mahal < cutoff)

#  	Delete the outliers.
noout = subset(iris1, mahal < cutoff)
str(noout)

```
iris$Species = as.factor(iris$Species)

```{r}
##Independent t-test: equal variances
t.test(Sepal.Length ~ Species,
 data = noout,
 var.equal = TRUE,
 paired = FALSE)

##independent t test unequal variances
t.test(Sepal.Length ~ Species,
 data = noout,
 var.equal = FALSE,
 paired = FALSE)

M = with(noout, tapply(Sepal.Length, Species, mean))
stdev = with(noout, tapply(Sepal.Length, Species, sd))
N = with(noout, tapply(Sepal.Length, Species, length))
M;stdev;N

```
Differences between groups was found: t(198)=-14.045,p=0.00000000000000022 in the Independent t-test set to equal variances. The results suggest that there is a small significant difference in the mean scores between the groups.The mean and standard deviation of setosa are both smaller than those of versicolor . According to statistics, the mean difference of = 5.986069 ,(df = 200) with a p-value < 0.05, so we reject our null hypothesis that there were no significant differences in Species and Sepal length results before the intervention.


